---
title: "vishnusagar"
output: html_document
---
# "Matrix, arrays and Data frames"
```{r}

print("Matrix, arrays and Data frames")

```

## 1.Matrix


```{r}
# In a matrix, numbers are arranged in a fixed number of rows and columns and usually, the numbers are the real numbers. *All the data elements must share a common basic type.

#`Funtion - matrix()`

```

```{r}
matrix ()
```
```{r}
mat = matrix (c(2,4,6,8,10,12),     # data or elements
nrow = 2,                        # no of rows first
ncol = 3,                       # no. of column second
byrow= TRUE)                     # true =[...    false = [..
                                  #       ...]            ..
                                  #                       ..]
mat

mat1 = matrix (c(2,4,6,8,10,12),
nrow = 2,
ncol = 3, 
byrow= FALSE)
mat1

# An element at the mth row and nth column 
mat [2,3]


#extract single element from a matrix
## To extract only the mth row of our matrix ‘mat’
mat[2,]
## To extract only the nth coln of our matrix ‘mat’
mat[,2]

```
###Creating R matrix through the arrangement of elements sequentially by row
```{r}             
#Creating R matrix through the arrangement of elements sequentially by row
# arrange_row =
arrang_row = matrix(c(4:15), nrow=4, byrow=TRUE)
print(arrang_row) 

```
###Creating R matrix by arranging elements sequentially by column
```{r}
#Creating R matrix by arranging elements sequentially by column
# arrang_col =
arrang_col = matrix(c(4:15), ncol = 4, byrow = FALSE)
arrang_col

```
#### Defining names of columns and rows in a matrix
```{r}
#Defining names of columns and rows in a matrix
# rows, cols
# dimnames = list(rows, cols)
# are the major codes required
# first Creating character vector of row names

rows = c("row1", "row2", "row3", "row4") #Creating  character vector of row names
cols = c("col 1", "col2", "col3")   #Creating  character vector of col names
mat5 = matrix(c(4:15), nrow=4, byrow=TRUE, dimnames=list(rows,cols))
mat5

rows = c("s1", "s2", "s3", "s4")
cols = c("pH", "EH", "EC", "TDS")
mat6= matrix(c(1:16), nrow=4, byrow = TRUE, dimnames=list(rows, cols))
mat6
```
#### Modify Matrix in R
```{r}
# Modify Matrix in R
#Assign a Single Element
mat6
mat6[2,2]
# replace single element 
mat6[2,2] =20
mat6
# replace column
mat6[,2] = c(21,23,25,27)     # mat[,x]= c(a1,a2,a3...an)    x is column no, c() is replacemnt
mat6
# replace row
mat6[3,] =c(20,30,40,50)
mat6


```
### Use of Relation Operators
```{r}
# Use of Relation Operators
# replacing elements with another elemnts
#mat[mat== element in matrix] = replacement element
mat6[mat6==50]=5
mat6
#Replacing elements that are less with 10 with 0
mat6[mat6 <10] = 0
mat6

```
### Addition of Rows and Columns
```{r} 
# first create a  new matrix
mat7 =matrix(c(1:12), nrow= 3, ncol =3, byrow=TRUE)
mat7
# add a column to our matrix using cbind() function , c()

mat7=cbind(mat7, c(10,11,12))
mat7
# add a row to our matrix using rbind() function, c()

mat7= rbind(mat7, c(20,30,40,50))
mat7

# dimension of the matrix  using the dim() function

dim(mat7)

# modify the dimension of the matrix  dim() = c(r,c)
# dim(mat7)= c(3,3)
# mat7

# transpose of the matrix using the t() function
t(mat7)
mat7

```
### R Matrix Operations
```{r}
#R Matrix Operations
#1. Addition
mat1 = matrix(c(1:8), nrow=4, ncol=4, byrow=TRUE)  #4x4 matrix
mat2= matrix(c(1:16), nrow=4, ncol=4, byrow=TRUE)   # 4*4
sum= mat1 + mat2 
print(sum)

# subtraction 
sub = mat1 -mat2 
print (sub)

# multiplication
prod= mat1 * mat2
print(prod)


#division

div = mat1/mat2
print (div)

# multiplication with a contstant
 prod = mat1*10
print (prod)

# multiplication with constant
prod= (mat1*mat2)* 10
print(prod)



```
## Array 

```{r}
## R Array 
#R arrays are the data objects which can store data in more than two dimensions. 
# In arrays, data is stored in the form of matrices, rows, and columns.
# 'array ()'
# `Array_NAME =  array(data, dim = (row_Size, column_Size, matrices, dimnames))`

array=array(data=NA, dim= length(data), dimname= NULL)
print(array)
```
#### Create Array in R

```{r}
# first Create two vectors of different lengths.
vector1= c(2,4,6,8)
vector2 = c(1,3,5,7,9,11,3,15)
# Take these vectors as input to the array.

array1 = array(c(vector1,vector2), dim = c(4,2,3))
array1

# Naming Columns And Rows
# first Create two vectors of different lengths.
vector1= c(2,4,6,8)
vector2 = c(1,3,5,7,9,11,3,15)
column.names= c("s1","s2","s3","s4")
row.names = c("L1","L2", "L3")
matrix.names= c("M1", "M2", "M3")
array2 = array(c(vector1,vector2), dim = c(3,4,3), dimnames= list(row.names,column.names,matrix.names))
array2

```

#Accessing R Array Elements
```{r}
# Print the mth row of the n th matrix of the array.
#print (array_name[m,,n])
# Print the 2nd row of the second matrix of the array.
print(array2[2,,2])

# Print the element in the 1st row and 3rd column of the 1st matrix.
print(array2[1,3,1])       

# Print the 2nd Matrix.
print(array2[,,2])

```
### Manipulating R Array Elements

```{r}
# Manipulating R Array Elements
#  first Create two vectors of different lengths
vector1= c(2,4,6)
vector2= c(1,3,5,7,9,11)
array1= array(c(vector1, vector2), dim= c(3,3,3))


vector3=c(12,14,16)
vector4=c(21,23,25,27,29,31)
array2= array(c(vector3, vector4), dim= c(3,3,3))

array1
array2

# create matrices from these array
matrix1 = array1[,,1]
matrix2 = array2[,,2]

# Add the matrices.

output = matrix1+matrix2
print(output)
```

## R Data Frame

```{r}
# The tabular data is referred by the data frames in which there are a number of observations(rows) or measurements (columns).
# ie, data table in excel like locations and parameters

# function for data frame
#`function - data.frame()`

# create a data frame

water_data= data.frame(
location_no= c(1:5),
parameter_names= c("pH", "EC", "EH", "Na", "Cl"),
surface_data = c(7,25,10,32,64),
Bottom_data  = c(8,36,12,42,72),
date_sampling = as.Date (c("2019-02-04", "2019-06-21", "2019-11-14", "2020-05-19","2020-09-25")),
stringsAsFactors = FALSE)

print(water_data)


# The structure of the data frame can see by using the str () function.
str(water_data)

# Extract Specific columns    $ for extraction

output= data.frame(water_data$surface_data,water_data$Bottom_data)
output


# Extract first two rows

output = water_data[1:2,]
output

# Expand R Data Frame - Add the column vector using a new column name.     $ for merging new data
water_data$quality_status = c("good","bad", "moderate","drinkable","contaminated")
water_data

water_data
#

```
### Bind the two data frames
```{r}
# Create the second R data frame    change the values for parametres
water_data2= data.frame(
location_no= c(6:8),
parameter_names= c("K", "Ba", "F"),
surface_data = c(32,56,52),
Bottom_data  = c(21,42,34),
date_sampling = as.Date (c("2019-02-04", "2019-06-21", "2019-11-14")),
quality_status = c("good","bad", "moderate"),
stringsAsFactors = FALSE)
water_data2

# binding the 2 data set use rbind()
# output = rbind(data1,data2)
output = rbind(water_data,water_data2)
output

 
```


## Including Plots

You can also embed plots, for example:

```{r pressure, echo=TRUE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
