---
title: "vishnusagar"
output: html_document
---
# "Matrix, arrays and Data frames"
```{r}

print("Matrix, arrays and Data frames")

```

## 1.Matrix


```{r}
# In a matrix, numbers are arranged in a fixed number of rows and columns and usually, the numbers are the real numbers. *All the data elements must share a common basic type.

#`Funtion - matrix()`

```

```{r}
matrix ()
```
```{r}
mat = matrix (c(2,4,6,8,10,12),     # data or elements
nrow = 2,                        # no of rows first
ncol = 3,                       # no. of column second
byrow= TRUE)                     # true =[...    false = [..
                                  #       ...]            ..
                                  #                       ..]
mat

mat1 = matrix (c(2,4,6,8,10,12),
nrow = 2,
ncol = 3, 
byrow= FALSE)
mat1

# An element at the mth row and nth column 
mat [2,3]


#extract single element from a matrix
## To extract only the mth row of our matrix ‘mat’
mat[2,]
## To extract only the nth coln of our matrix ‘mat’
mat[,2]

```
###Creating R matrix through the arrangement of elements sequentially by row
```{r}             
#Creating R matrix through the arrangement of elements sequentially by row
# arrange_row =
arrang_row = matrix(c(4:15), nrow=4, byrow=TRUE)
print(arrang_row) 

```
###Creating R matrix by arranging elements sequentially by column
```{r}
#Creating R matrix by arranging elements sequentially by column
# arrang_col =
arrang_col = matrix(c(4:15), ncol = 4, byrow = FALSE)
arrang_col

```
#### Defining names of columns and rows in a matrix
```{r}
#Defining names of columns and rows in a matrix
# rows, cols
# dimnames = list(rows, cols)
# are the major codes required
# first Creating character vector of row names

rows = c("row1", "row2", "row3", "row4") #Creating  character vector of row names
cols = c("col 1", "col2", "col3")   #Creating  character vector of col names
mat5 = matrix(c(4:15), nrow=4, byrow=TRUE, dimnames=list(rows,cols))
mat5

rows = c("s1", "s2", "s3", "s4")
cols = c("pH", "EH", "EC", "TDS")
mat6= matrix(c(1:16), nrow=4, byrow = TRUE, dimnames=list(rows, cols))
mat6
```
#### Modify Matrix in R
```{r}
# Modify Matrix in R
#Assign a Single Element
mat6
mat6[2,2]
# replace single element 
mat6[2,2] =20
mat6
# replace column
mat6[,2] = c(21,23,25,27)     # mat[,x]= c(a1,a2,a3...an)    x is column no, c() is replacemnt
mat6
# replace row
mat6[3,] =c(20,30,40,50)
mat6


```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=TRUE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
